### 2016-12-27 paper.tex
...

\documentclass{article}
%\documentclass[letterpaper, 10 pt, conference]{ieeeconf}

%\IEEEoverridecommandlockouts        % This command is only needed if you want to use the \thanks command
%\overrideIEEEmargins                        % See the \addtolength command later in the file to balance the column lengths

% Name clashes workaround
%\makeatletter
%\let\IEEEproof\proof
%\let\IEEEendproof\endproo f
%\let\proof\@undefined
%\let\endproof\@undefined
%\makeatother

\usepackage{epsfig} 
\usepackage{amsmath} 
\usepackage{amssymb} 
\usepackage{amsthm}  
\usepackage{hyperref}
\usepackage{tikz}
\usetikzlibrary{automata,arrows,decorations.pathmorphing,decorations.markings,positioning}
\usetikzlibrary{shapes}
%\usepackage{balance}
\usepackage{scrextend}
\addtokomafont{labelinglabel}{\sffamily}

% Theorem like env.
\newtheorem{theorem}{Theorem}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{claim}[theorem]{Claim}
\newtheorem{lemma}[theorem]{Lemma}

\theoremstyle{definition}
\newtheorem{definition}[theorem]{Definition}     
\newtheorem{conjecture}[theorem]{Conjecture}
\newtheorem{observation}[theorem]{Observation}
\newtheorem{example}[theorem]{Example}
\newtheorem{remark}[theorem]{Remark}
\newtheorem{corrolary}[theorem]{Corrolary}

%\newcommand{\RotateLeft}{rlft}
\DeclareMathOperator{\RotateLeft}{RotLft}

%------------------------------------------- preamble ---------------------------

\newcommand{\R}{{\mathbb{R}}}
\newcommand{\N}{{\mathbb{N}}}
\newcommand{\Z}{{\mathbb{Z}}}

\newcommand{\REF}[2]{#1~\ref{#2}}

\newcommand{\T}[1]{\langle{#1}\rangle}
\newcommand{\rr}[2]{R^{#2}({#1})}
\newcommand{\rl}[2]{R^{-{#2}}({#1})}

\newcommand{\AorB}[2]{\bigl(#2\text{?}A(#1)\text{:}B(#1)\bigr)}

\usepackage{amssymb} 
\usepackage{algpseudocode,algorithm}
\usepackage{subfigure}

\newcommand{\nega}   [1]{\overline{#1}}
\newcommand{\companion}{comp}
\newcommand{\shlsame}[1]{shl\text{-}same(#1)}
\newcommand{\shlopp} [1]{shl\text{-}opp(#1)}
\newcommand{\shlone} [1]{shl\text{-}one(#1)}
\newcommand{\shlzero}[1]{shl\text{-}zero(#1)}
\newcommand{\idx}    [1]{idx(#1)}

%\pagestyle{plain}

\title{An Efficient Construction of the Prefer-Max Sequence and of an Infinite de Bruijn Sequence}

\author{}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}
\maketitle

\section{Introduction}

A $k$-ary de Bruijn sequence of span $n$ is a cyclic sequence of length $k^n$ in which each $k$-ary string of length $n$ appears exactly once as
a substring. A well known construction of such a sequence was proposed by Ford ~\cite{Ford57} as follows:

\begin{algorithm}
Repeatedly append the symbol $k-1$ if the word formed by the $n$ most recent symbols (padded with zeros on the left, if not enough symbols exist) is new, otherwise append $k-2$, otherwise $k-3$, etc. Stop after appending $n$ consecutive zeros. 
\caption{The prefer-max construction.}
\end{algorithm}

For example, if we choose $k=3$ and $n=3$ we get the sequence:
$$\langle2,2,2,1,2,2,0,2,1,1,2,1,0,2,0,1,2,0,0,1,1,1,0,1,0,0,0\rangle.$$
Note that this sequence begins with a prefix (the first $19$ symbols) that contains all the words in $\{0,1,2\}^n \setminus \{0,1\}^n$ as substrings. Note also that the rest of the sequence, after this prefix, is exactly the sequence that we would have obtained if we ran the construction with the alphabet $k=2$ (and the same $n$). The first result of this paper, that we call the Onion lemma, is that this is not a coincidence:

\begin{theorem}[Onion]
	For any $k,n \in \N$, the prefix of length $k^n-(k-1)^n$ of the sequence generated by the prefer-max construction is such that its $n$-length substrings are exactly $\{0,\dots,k\}^n \setminus \{0,\dots,k-1\}^n$, i.e., all the words of length $n$ that contain the symbol $k$. Furthermore, the suffix that begins with the 
	$(k^n-(k-1)^n+1)$th symbols is exactly the sequence generated by the prefer-max construction with the parameters $k-1$ and $n$.
	\label{thm:onion}
\end{theorem}

This observation lead us to look at the reverse of the prefer-max sequence. Specifically, the structure identified in Theorem~\ref{thm:onion} means that if we somehow manage to construct the sequence backwards, we do not have to specify $k$ ahead of time because we can go on indefenitiley while $k$ grows in layers: we first write $0^n$, then a word whose substrings of length $n$ are $\{0,1\}^n$, then extend to $\{0,1,2\}^n$, and so on.

The second contribution of this work is then an efficient construction of this reversed sequence. This is formulated in Algorithm~\ref{inf-construction}. We say that this construction is efficient because each step can be computed in $O(n^2)$ operations by enumerating all the $n$ rotations of $\sigma w$ and by comparing them lexicographically (reading each word from right to left).

\begin{algorithm}[!h]
	Append $n$ zeros. Then, repeatedly, let $\sigma w$ be the word formed by the $n$ most recent symbols. Append $\sigma+1$ if $(\sigma+1)w$ is of the form $u0^l$ where $0^lu$ is maximal in right-to-left lexicographic order among all of its rotations. If $(\sigma+1)w$ is not of that form and $\sigma w$ is, append the symbol $0$. Otherwise, if neither $(\sigma+1)w$ nor $(\sigma)w$ are of that form, append $\sigma$.
	\caption{An infinite de Bruijn sequence.}
	\label{inf-construction}
\end{algorithm}

The fact that this generates the reverse of the prefer-max sequence is formulated in the following theorem:

\begin{theorem} 
The first $k^n$ symbols generated by Algorithm~\ref{inf-construction} form the reverse of the sequence generated by the prefer-max construction with the parameters $k$ and $n$.	
\end{theorem}

In particular, the above theorem says that the sequence generated by Algorithm~\ref{inf-construction} is a de Bruijn sequence, i.e., that each word in $\N^n$ appears exactly once as a substring.

While our main focus is on analyzing the reversed sequence, we also have a result about the prefer-max sequence itself. Our result gives the first (to the best of our knowledge) efficient construction for that sequence:

\begin{algorithm}
	Append $n-1$ zeros and then the symbol $k$. Then, repeatedly, let $\sigma w$ be the word formed by the $n$ most recent symbols. Append the symbol $\sigma-1$  if $\sigma w$  is of the form specified in Algorithm~\ref{inf-construction}. Otherwise, if there is a maximal $\sigma'$ such that $\sigma'w$ is of that form, append the symbol $\sigma'$. Otherwise, if there is no such $\sigma'$, append the symbol $\sigma$.

	\caption{An efficient construction of the prefer-max sequence.}
	\label{efficient-pref-max}
\end{algorithm}

Each step of Algorithm~\ref{efficient-pref-max} can be computed in $O(n^2)$ time because it suffices to examine for $\sigma'$ only symbols that are $\sigma w$ or are smaller by one than a symbol in $\sigma w$, as we will establish later in the paper.
The fact that this is indeed the prefer-max sequence is stated in the following theorem:

\begin{theorem}
	The sequence generated by Algorithm~\ref{efficient-pref-max} is the prefer-max sequence with parameters $k$ and $n$.
\end{theorem}

A central tool to our analysis of the sequences descibed above is an alternative formulation of the Ford sequence. For this, we need to introduce some language, as follows.

The $n$ dimensional infinite de Bruijn graph is the graph $G_{n}=\T{V,E}$ where $V=\N^n$ and $E=\{ \T{\sigma_1 w, w \sigma_2} \colon \sigma_1,\sigma_2\in\N, w \in \N^{n-1}\}$. The de Bruijn sequences, defined in the beginning of this paper, are in one-to-one correspondence with the Hamiltonian cycles in the de Bruijn graph, as follows: (1) Concatenating the first letter of each vertex in a cycle produces a de Bruijn sequence of order $n$; and (2) A cycle can be constructed from a de Bruijn sequence $\T{\sigma_1,\dots,\sigma_{2^n}}$ of order $n$ by visiting the vertex $\sigma_1\cdots\sigma_n$, then $\sigma_2\cdots\sigma_{n+1}$ and so on.

Let $rtl(\sigma_1\sigma_2\dots\sigma_n) = \sigma_2\dots\sigma_n\sigma_1$ denote the rotation of a word to the left and let $rtl^i(w)$ denote a rotation of $i$ letters to the left. A \emph{necklace} in the de Bruijn graph $G_n$ is the set $\{rtl^i(w)\}_{i=1}^n$ where $w \in \N^n$, i.e., it is an equivalence class of $n$-character strings over $\N$ when strings that are rotations of each other are considered equivalent. A $w\in\N^n$ is called a \emph{key word} if it is larger in right-to-left lexicographic order than all of its rotations (all the words in its necklace).  Note the resemblance of this to the notion of Lyndon words~\cite{??}, on which we will elaborate later.

We are now ready to introduce another construction:

\begin{algorithm}[!h]
	Start with the one element sequence $\T{0^n}$. The sequence at the $i$th step is constructed by adding words to the previous sequence as follows: Let $k_i$ be the $i$th key word in right-to-left lexicographic order. Let $l$ be the number of leading zeros in $k_i$ ($l$ may be zero) and let $\sigma \in \N$ and $w\in\N^{n-l-1}$ be such that $k_i=0^l(\sigma+1)w$. Now, add the sub-sequence $\T{rtl^i(w0^l(\sigma+1))}_{i=0}^{n-1}$ between the word $\sigma w 0^l$ and its follower in the previously constructed sequence.
	\caption{A necklace joining construction.}
	\label{necklace-joining}
\end{algorithm}

To following proposition establishes that the above construction describes an infinite sequence of words:

\begin{proposition}
	For every $l>0$ there is $m>0$ such that for every $i>m$ the insertions of words at the $i$th step of Algorithm~\ref{necklace-joining} are all after the $l$th element of the sequence.
\end{proposition}

Let $S=\T{w_i}_{i=0}^\infty$ be the infinite sequence all whose prefixes agree with infinitely many prefixes of the finite sequences generated at the steps of Algorithm~\ref{necklace-joining}. Since every word is a rotation of some key word, we have that this sequence contains all the words in $\N^n$. Since there is a directed edge in $G_n$ from each word in this sequence to its follower, the sequence corresponds to an Hamiltonian path in $G_n$. The following theorem states that this path maps, by the standard correspondence described above, to the reverse of the prefer-map sequence:

\begin{theorem}
	If we take the first letter of each word in $S$ we get the sequence described in Algorithm~\ref{inf-construction}.
\end{theorem}

The

\end{document}